IPv6 over Low-Power Wireless Personal Area Networks\\ MSc. Thesis Proposal

### 2009-12-07 trunk/doc/msc-proposal/msc-proposal.tex
proposal: HC2

%%
%	Introduction
%%
\chapter{Introduction}
* There are a lot of applications where we need many small devices + examples; why it's important;
* LoWPAN\\
* The purpose of the proposed work will be to investigate smth. The proposal describes problems associated with enabling IP communication with devices in a LoWPAN.\\
* As preparation for this work, several related areas have been studied in more
depth and the collected background information; Structure of the proposal.\\

@@

\section{Adaptation layer}\label{sec:adapt.layer}
The IPv6 minimum Maximum Transmission Unit (MTU) size is defined as 1280 octets. As described in Chapter \ref{ch:ieee802.15.4}, the maximum IEEE 802.15.4 frame size is limited to 127 octets.  Taking into account the maximum frame overhead of 25 octets, 102 octets are left at the media access control layer. Link-layer security imposes further overhead, which in the maximum case (21 octets for AES-CCM-128) leaves only 81 octets available, which is far below the minimum IPv6 packet size. Therefor, a fragmentation and reassembly adaptation layer is provided at the layer below IP.

@@

\subsection{Broadcast Header}\label{subsec:broad.header}
Broadcast is an additional mesh routing functionality. For that a broadcast header is defined the format of which is shown in Figure \ref{fig:broad.header}.
\begin{figure}[htp]
\centering
\begin{mylisting}
\begin{verbatim}
                       1
   0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5
  +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
  |0|1|0|1|0|0|0|0|Sequence Number|
  +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{mylisting}
\caption{Broadcast Header
}\label{fig:broad.header}
\end{figure}

@@

\section{Header compression}
As described in Section \ref{sec:adapt.layer}, 81 octets are left in a IEEE 802.15.4 frame for IPv6. The IPv6 header is 40 octets long which leaves only 41 octets for upper-layer protocols, like UDP.  The latter uses 8 octets in the header, which leaves 33 octets for application data. 
The fragmentation and reassembly layer, described in Section \ref{sec:adapt.layer}, will also use at least one additional octet, what makes header compression almost being unavoidable.

@@

%% Bibliography
\newpage
\bibliographystyle{plain}
\bibliography{msc-proposal}
\addcontentsline{toc}{chapter}{Bibliography}

\nocite{ieee802.15.4} 
\nocite{eui64} 
\nocite{rfc2460}
\nocite{rfc2464}
\nocite{rfc4291} 
\nocite{rfc4443}
\nocite{rfc4861}
\nocite{rfc4862}
\nocite{rfc4919}
\nocite{rfc4944} 
\nocite{draft-usecases-05} 
\nocite{draft-hc-06} 
\nocite{draft-nd-06}

>>>>

%%
%	Introduction
%%
\chapter{Introduction}
* There are a lot of applications where we need many small devices + examples; why it's important;
* LoWPAN\\
* The purpose of the proposed work will be to investigate something. The proposal describes problems associated with enabling IP communication with devices in a LoWPAN.\\
* As preparation for this work, several related areas have been studied in more
depth and the collected background information; Structure of the proposal.\\

@@

\section{Adaptation Layer}\label{sec:adapt.layer}
The IPv6 minimum Maximum Transmission Unit (MTU) size is defined as 1280 octets. As described in Chapter \ref{ch:ieee802.15.4}, the maximum IEEE 802.15.4 frame size is limited to 127 octets.  Taking into account the maximum frame overhead of 25 octets, 102 octets are left at the media access control layer. Link-layer security imposes further overhead, which in the maximum case (21 octets for AES-CCM-128) leaves only 81 octets available, which is far below the minimum IPv6 packet size. Therefor, a fragmentation and reassembly adaptation layer is provided at the layer below IP.

@@

\subsection{Broadcast Header}\label{subsec:broad.header}
Broadcast is an additional mesh routing functionality. For that a broadcast header is defined the format of which is shown in Figure \ref{fig:broad.header}.
\begin{figure}[htp]
\begin{mylisting}
\begin{verbatim}
                                       1
                   0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5
                  +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
                  |0|1|0|1|0|0|0|0|Sequence Number|
                  +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{mylisting}
\caption{Broadcast Header
}\label{fig:broad.header}
\end{figure}

@@

\section{Header Compression}
As described in Section \ref{sec:adapt.layer}, 81 octets are left in a IEEE 802.15.4 frame for IPv6. The IPv6 header is 40 octets long which leaves only 41 octets for upper-layer protocols, like UDP.  The latter uses 8 octets in the header, which leaves 33 octets for application data. 
The fragmentation and reassembly layer, described in Section \ref{sec:adapt.layer}, will also use at least one additional octet, what makes header compression almost being unavoidable.

RFC 4944 \cite{rfc4944} defines stateless header compression mechanisms for IPv6 datagrams (LOWPAN\_HC1 and LOWPAN\_HC2) to reduce the relatively large IPv6 and UDP headers. However, LOWPAN\_HC1 and LOWPAN\_HC2 are insufficient for most practical uses of 6LoWPAN networks, since they are effective only for link-local unicast communication, where IPv6 addresses derived directly from IEEE 802.15.4 addresses. When communicating with devices external to the LoWPAN or in a route-over configuration where IP forwarding occurs within the LoWPAN, LOWPAN\_HC1 carries both IPv6 source and destination addresses in-line. Internet Draft document \cite{draft-hc-06} defines an encoding format, LOWPAN\_IPHC, for effective compression of unique local, global, and multicast IPv6 addresses based on shared contexts. In addition, it also defines an encoding format, LOWPAN\_NHC, for arbitrary next header compression.

\subsection{HC1 Header Encoding}
The HC1 compression format does not use any shared context and compresses headers relying on information pertaining to the entire link. The following IPv6 header values are expected to be common on 6LoWPAN network: Version is IPv6; both IPv6 source and destination addresses are link local;  the IPv6 interface identifiers for the source or destination addresses can be inferred from the layer two source and destination addresses; the packet length can be inferred either from layer two or from the "datagram\_size" field in the fragment header; both the Traffic Class and the Flow Label are zero; and the Next Header is UDP, ICMP or TCP. The Hop Limit (8 bits) field is the only one that always needs to be carried. Using the HC1 encoding allows to compress a 40 byte IPv6 header to 2 bytes.

The HC1 encoding uses a dispatch value of $000010$ to specify that the payload following the dispatch header, described in Section \ref{subsec:dispatch.header}, is a HC1 compressed IPv6 header. The HC1 header is 8 bits. Bits 0 and 1 specify the encoding for the IPv6 source prefix and interface identifier respectively: $0$ -- the field is carried in-line; $1$ -- link-local address is assumed. Bits 2 and 3 specify how the destination address is encoded (the same encoding as for the source address). If bit 4 is set then  Traffic Class and Flow Label fields are zero, otherwise their values are not compressed and carried in-line. Bits 5 and 6 indicate the Next Header: $00$ -- the field is not compressed and carried inlie, $01$ corresponds to UDP, $10$ -- ICMP, $11$ -- TCP. The last bit specifies whether HC1 header is immediately followed by more header compression bits of the HC2 encoding format.

By using this encoding an IPv6 header can be compressed to different degrees and only non-compressed fields need to be sent. The only field that is always present is the Hop Limit (8 bits). It follows the HC1 encoding fields. Other non-compressed fields follow the Hop Limit.

\subsection{HC\_UDP Header Encoding}
The HC\_UDP encoding is used for compression of a UDP header. It is applied only if bits $5$ and $6$ in HC1 specify that the protocol that follows the IPv6 header is UDP and bit 7 indicates that the HC2 encoding is used. The HC\_UDP encoding allows to compress the UDP header to 4 octets instead of the original 8 octets.

The first $8$ bits of the HC\_UDP encoding specify how the UDP header is compressed. Both the source and destination port fields may be compressed to $4$ bits. It is indicated by bits $0$ and $1$ of the HC\_UDP encoding for the source and destination ports respectively. In such case, the actual $16$-bit port is obtained by calculating $61616$ + short port value, which is a $4$ bit value carried in-line. If bit $2$ of the HC\_UDP encoding is set then the UDP length header field is computed from IPv6 header length information, otherwise it's carried in-line. HC\_UDP encoding bits 3-7 are reserved for future use. The UDP header's checksum field is not compressed and is therefore carried in full.

The non-compressed or partially compressed fields in the UDP header always follow the IPv6 header. All present UDP header in-line fields always appear in the same order as the corresponding fields appear in a normal UDP header \cite{rfc780}, e.g., source port, destination port, length, and checksum.

\subsection{IPHC Header Encoding}

\subsection{NHC Header Encoding}

@@

% Bibliography
\newpage
\bibliographystyle{plain}
\bibliography{msc-proposal}
\addcontentsline{toc}{chapter}{Bibliography}

\nocite{ieee802.15.4} 
\nocite{eui64} 
\nocite{rfc780}
\nocite{rfc2460}
\nocite{rfc2464}
\nocite{rfc4291} 
\nocite{rfc4443}
\nocite{rfc4861}
\nocite{rfc4862}
\nocite{rfc4919}
\nocite{rfc4944} 
\nocite{draft-usecases-05} 
\nocite{draft-hc-06} 
\nocite{draft-nd-06}
